---
title: "Project 1"
author: "Miguel Toralla"
date: "21/11/2020"
output:
  pdf_document: default
  html_document: default
  md_document: 
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preprocessing
First, we need to unzip and load the data into R. This next chuck of code will be cached so that we will not be loading the data over and over again.

```{r unzip, cache=TRUE}
unzip(zipfile = "repdata_data_activity.zip")
data <-read.csv(file = "activity.csv", header = TRUE, na.strings = "NA", colClasses = c("numeric", "Date", "numeric"))
data$date <- as.Date(data$date, "%Y-%m-%d")
head(data)
summary(data)
```
## What is mean total number of steps taken per day?

Now, we are going to make a histogram of the total number of steps taken each day. To do this, we will be using the *ggplot2* package. Also, we will group total steps by day, and consider only the complete cases. 

```{r histogram data, results="hide", message=FALSE}
library(ggplot2)
library(dplyr)
data_histogram <- data[complete.cases(data),] %>% group_by(date)%>%
        summarise(steps= sum(steps, na.rm = TRUE))
```

```{r plot,message=FALSE}
ggplot(data_histogram, aes(steps)) + 
        geom_histogram()
```

Now that we have the steps taken by day, we can easily compute the mean and the median.
```{r mean and median}
mean(data_histogram$steps)
median(data_histogram$steps)
```
## What is the average daily activity pattern?

We are going to make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days . To do this, we need to group the data by time interval, and compute the mean per interval.

```{r, message=FALSE}
data_timeSeries <- data %>% group_by(interval) %>%
        summarise(daily_average = mean(steps, na.rm = TRUE))


ggplot(data_timeSeries, aes(interval, daily_average)) + geom_line()
```

Now, we can easily obtain the maximum average value and its respective interval:


```{r}

data_timeSeries[which.max(data_timeSeries$daily_average),]
```

## Imputing missing values

We need to check how many rows have missing values. 

```{r}
sum(!complete.cases(data))
```
Now, we want to fill all those NA values. For this, we will use the daily average for the respective time interval. We will check if a value is missing and if it is, we will replace fill it. 

```{r}
filled_data <- data
for(i in 1:length(filled_data$steps))
      if(is.na(filled_data$steps[i]))
            filled_data$steps[i] <- data_timeSeries$daily_average[i]
```

With this done, we can remake the histogram we make before.
```{r, results="hide", message=FALSE}
data_histogram2 <- filled_data %>% group_by(date)%>%
        summarise(steps= sum(steps, na.rm = TRUE))
```

```{r, message=FALSE}
ggplot(data_histogram2, aes(steps)) + 
        geom_histogram()
```

The new mean and median for this data set is: 

```{r}
mean(data_histogram2$steps)
median(data_histogram2$steps)
```

From this, we see that the mean and the median differ from the once we computed before, being the newer onces smalles. 


## Are there differences in activity patterns between weekdays and weekends?

To find out if there are any differences in the activity between weekends and weekdays, we will create a new data set with a new factor variable with 2 levels: weekday and weekend. 

```{r}
days <-c("lunes","martes","miércoles","jueves","viernes")
week_activity <- filled_data
week_activity$activity <- ifelse(weekdays(week_activity$date) %in% days , "weekday" ,"weekend")

```

At last, we want to make a plot comparing the weekday activity and the weekend activity.

```{r, message=FALSE}
data_timeSeries2 <- week_activity %>% group_by(activity, interval) %>%
        summarise(daily_average = mean(steps, na.rm = TRUE))


ggplot(data_timeSeries2, aes(interval, daily_average)) + geom_line() + facet_grid(activity~.)
```